---
author: Raven Garvey
title: Current and potential roles for archaeology in the development of cultural evolutionary theory (published inPhilosophical Transactions B)
date: 17 November, 2017
output: bookdown::pdf_document2
---

# Introduction

Code for research published in:

Garvey Raven Current and potential roles of archaeology in the development of cultural evolutionary theory. _Phil. Trans. R. Soc. B ans. R. Soc. B_ http://doi.org/10.1098/rstb.2017.0057

These electronic supplementary materials are also available online at https://dx.doi.org/10.6084/m9.figshare.c.3965853.

```{r}
data <- read.csv("data/data.csv", skip = 6)
str(data)
```

# Multi panel Washita figure

## Create custom colors

```{r}
makeTransparent <- function(black, alpha = 200){
	newColor <- col2rgb(black)
	apply(newColor, 2, function(curcoldata)			
	     {rgb(red = curcoldata[1], 
	          green = curcoldata[2],
	          blue = curcoldata[3],
	          alpha = alpha, maxColorValue = 255)})
}
tBlack   <- makeTransparent("gray40")
tOrange  <- makeTransparent("orange")
tBlue    <- makeTransparent("cornflowerblue")
```

## Plot and simulation parameters

Make data frame

```{r}
D <- data.frame(data$wash_width, data$wash_basewidth,
                data$wash_neck, data$wash_haftlength,
                data$wash_length, data$wash_midlength, 
                data$wash_thick, data$wash_bladelength, 
                data$wash_weight)
```

Set counter to # variables (i.e., # figure rows)

```{r}
count <- c(1:9)       
```

Set the attribute-specific n.measures

```{r}
n.measures <- c(162, 157,
                249, 227,
                136, 136,
                259, 135,
                87)
```

y-axis panel plot labels

```{r}
labs <- c("width\n(mm)", "base width\n(mm)",
          "neck width\n(mm)", "haft length\n(mm)",
          "length\n(mm)", "mid length\n(mm)",  
          "thickness\n(mm)", "blade length\n(mm)",
          "weight\n(g)")
```

Simulation parameters

```{r}
set.seed(1212)            # set seed
n.houses       <- 100     # set number of houses at site
y              <- 1000    # set number of simulation runs

## plotting parameters
col.density.1  <- tBlack     # set density color for 10%
col.outline.1  <- tBlack     # set density outline color 10%
col.density.05  <- tOrange   # set density color for 5%
col.outline.05  <- tOrange   # set density outline color 5%
col.density.03  <- tBlue     # set density color for 3%
col.outline.03  <- tBlue     # set density outline color 3%
col.mean       <- "black"    # set col for mean of arch sample
cex.ylab       <- 1.2        # set y-axis label size
cex.ylabtick   <- 1.1        # set y-axis tick label size
cex.xlabtick   <- 1.1        # set x-axis tick label size

x.lim2   <- c(6, 7,         # set x limit for each plot
              3.5, 4,       # (this also affects tick mark placement)
              50, 40,
              1.4, 35,
              0.12)

y.lim2   <- c(1.2, 1,       # set y limit for each plot
              1.75, 1.2,    
              0.15, 0.15,
              5, 0.15,
              45)
```

## The plot

```{r}
pdf(file="figures/Garvey_Figure3a.pdf", 
    height = 12, width = 12)

par(mfcol = c(length(count), 3))
par(oma = c (6, 10, 5, 2))

# column 1 (SD = 0.1)

diff10     <- NULL  #create empty vector to hold calculate diff between
                    #simulated population mean and observed mean (standardized)


for(x in count) {   # looping procedure for each plotting pannel

SD       <- 0.1    # set standard deviation multiplier for random draws 
d        <- na.omit(D[, x])     # select the data set to sample
N        <- n.measures[x]       # select the appropriate n.measures

metadig   <- NULL  #create empty vector to hold digs
var.Kdigs <- NULL  #create empty vector to hold variances
cv.Kdigs  <- NULL  #create empty vector to hold CVs

for(i in 1:y) {    # "master" loop that  repeats
                   # "archaeological sampling"" from 
                   # "generational learning sample"" one thousand times

    all.houses <-  NULL         # create an empty vector to hold results
    
    for(j in 1:n.houses) {       #loop through all houses       

               house.seed <- sample(d, 1, replace = TRUE)   
               house.g1 <- rnorm(4, mean = house.seed, 
                                 sd = SD * mean(house.seed)) 
               house.g2 <- rnorm(4, mean = mean(house.g1), 
                                 sd = SD * mean(house.g1)) 
               house.g3 <- rnorm(4, mean = mean(house.g2), 
                                 sd = SD * mean(house.g2))
               house.g4 <- rnorm(4, mean = mean(house.g3), 
                                 sd = SD * mean(house.g3))
               total.house.sample <- as.vector(c(house.g1,
                                                  house.g2,
                                                  house.g3,
                                                  house.g4))
           all.houses <- c(total.house.sample, all.houses)
          
        }
           dig <- sample(all.houses, N, replace = TRUE)
           metadig <- c(metadig, dig)
           var.Kdigs <- c(var(dig), var.Kdigs)
           cv.Kdigs <- c(sd(dig)/mean(dig), cv.Kdigs)
}

diff10 <- c(diff10, ((var(d)-mean(var.Kdigs))/sd(var.Kdigs)))

#plotting
dp.md <- density(var.Kdigs)
par(mai = c(0.25, 0.5, 0.1, 0))


plot(dp.md, col = col.outline.1, bty = 'n',
     main = "", xlab = "", ylab = "",
     xlim = c (0, x.lim2[x]),
     ylim = c (0, y.lim2[x]),
     xaxt = "n", yaxt = "n")

axis(1, at = c(0, 0.25 * x.lim2[x],
               0.5 * x.lim2[x], 0.75 * x.lim2[x],
               x.lim2[x]),
     cex = cex.xlabtick)

# alternative, with only 3 x-axis ticks
# axis(1, at = c(0, 0.5 * x.lim2[x],
#                x.lim2[x]),
#      cex = cex.xlabtick)

axis(2, at = c(0, 0.5 * y.lim2[x],
               y.lim2[x]),
     line = 1.5, las = 2,
     cex = cex.ylabtick)

polygon(dp.md, col = col.density.1)
abline(v = var(d), col = col.mean, lwd = 2)
mtext (text = labs[x], side = 2, 
       las = 1, line = 5.5,
       cex = 1.2, outer = FALSE)

}

mtext (text = "A. CV = 10%", side = 3, 
       las = 1, line = 73,
       cex = 1.5, outer = FALSE)


# column 2 (CV = 5%)

diff5     <- NULL  #create empty vector to hold calculate diff between
                   #simulated population mean and observed mean (standardized)


for(x in count) {   # looping procedure for each plotting pannel

SD       <- 0.05    # set standard deviation multiplier for random draws 
d        <- na.omit(D[, x])     # select the data set to sample
N        <- n.measures[x]       # select the appropriate n.measures

metadig   <- NULL  #create empty vector to hold digs
var.Kdigs <- NULL  #create empty vector to hold variances
cv.Kdigs  <- NULL  #create empty vector to hold CVs

for(i in 1:y) {    # "master" loop that should repeat
                   # "archaeological sampling"" from 
                   # "generational learning sample"" one thousand times

    all.houses <-  NULL         # create an empty vector to hold results
    
    for(j in 1:n.houses) {       #loop through all houses       

               house.seed <- sample(d, 1, replace = TRUE)   
               house.g1 <- rnorm(4, mean = house.seed, 
                                 sd = SD * mean(house.seed)) 
               house.g2 <- rnorm(4, mean = mean(house.g1), 
                                 sd = SD * mean(house.g1)) 
               house.g3 <- rnorm(4, mean = mean(house.g2), 
                                 sd = SD * mean(house.g2))
               house.g4 <- rnorm(4, mean = mean(house.g3), 
                                 sd = SD * mean(house.g3))
               total.house.sample <- as.vector(c(house.g1,
                                                  house.g2,
                                                  house.g3,
                                                  house.g4))
           all.houses <- c(total.house.sample, all.houses)
          
        }
           dig <- sample(all.houses, N, replace = TRUE)
           metadig <- c(metadig, dig)
           var.Kdigs <- c(var(dig), var.Kdigs)
           cv.Kdigs <- c(sd(dig)/mean(dig), cv.Kdigs)
}


diff5 <- c(diff5, ((var(d)-mean(var.Kdigs))/sd(var.Kdigs)))


#plotting
dp.md <- density(var.Kdigs)
par(mai = c(0.25, 0.5, 0.1, 0))

plot(dp.md, col = col.outline.05, bty = 'n',
     main = "", xlab = "", ylab = "",
     xlim = c (0, x.lim2[x]),
     ylim = c (0, y.lim2[x]),
     xaxt = "n", yaxt = "n")

axis(1, at = c(0, 0.25 * x.lim2[x],
               0.5 * x.lim2[x], 0.75 * x.lim2[x],
               x.lim2[x]),
     cex = cex.xlabtick)


polygon(dp.md, col = col.density.05)
abline(v = var(d), col = col.mean, lwd = 2)

}

mtext (text = "B. CV = 5%", side = 3, 
       las = 1, line = 73,
       cex = 1.5, outer = FALSE)


# column 3 (CV = 3%)

diff3     <- NULL  #create empty vector to hold calculate diff between
                   #simulated population mean and observed mean (standardized)



for(x in count) {   # looping procedure for each plotting pannel

SD       <- 0.03    # set standard deviation multiplier for random draws 
d        <- na.omit(D[, x])     # select the data set to sample
N        <- n.measures[x]       # select the appropriate n.measures

metadig   <- NULL  #create empty vector to hold digs
var.Kdigs <- NULL  #create empty vector to hold variances
cv.Kdigs  <- NULL  #create empty vector to hold CVs

for(i in 1:y) {    # "master" loop that should repeat
                   # "archaeological sampling"" from 
                   # "generational learning sample"" one thousand times

    all.houses <-  NULL         # create an empty vector to hold results
    
    for(j in 1:n.houses) {       #loop through all houses       

               house.seed <- sample(d, 1, replace = TRUE)   
               house.g1 <- rnorm(4, mean = house.seed, 
                                 sd = SD * mean(house.seed)) 
               house.g2 <- rnorm(4, mean = mean(house.g1), 
                                 sd = SD * mean(house.g1)) 
               house.g3 <- rnorm(4, mean = mean(house.g2), 
                                 sd = SD * mean(house.g2))
               house.g4 <- rnorm(4, mean = mean(house.g3), 
                                 sd = SD * mean(house.g3))
               total.house.sample <- as.vector(c(house.g1,
                                                  house.g2,
                                                  house.g3,
                                                  house.g4))
           all.houses <- c(total.house.sample, all.houses)
          
        }
           dig <- sample(all.houses, N, replace = TRUE)
           metadig <- c(metadig, dig)
           var.Kdigs <- c(var(dig), var.Kdigs)
           cv.Kdigs <- c(sd(dig)/mean(dig), cv.Kdigs)
}


diff3 <- c(diff3, ((var(d)-mean(var.Kdigs))/sd(var.Kdigs)))


#plotting
dp.md <- density(var.Kdigs)
par(mai = c(0.25, 0.5, 0.1, 0))

plot(dp.md, col = col.outline.03, bty = 'n',
     main = "", xlab = "", ylab = "",
     xlim = c (0, x.lim2[x]),
     ylim = c (0, y.lim2[x]),
     xaxt = "n", yaxt = "n")

axis(1, at = c(0, 0.25 * x.lim2[x],
               0.5 * x.lim2[x], 0.75 * x.lim2[x],
               x.lim2[x]),
     cex = cex.xlabtick)

polygon(dp.md, col = col.density.03)
abline(v = var(d), col = col.mean, lwd = 2)

}

mtext (text = "C. CV = 3%", side = 3, 
       las = 1, line = 73,
       cex = 1.5, outer = FALSE)


dev.off()
```

```{r}
knitr::include_graphics("figures/Garvey_Figure3a.pdf")
```

## Kernel density plot to compare simulation results

```{r}

pdf(file="figures/Garvey_Figure3b.pdf", 
    height = 3, width = 6)

d10 <- density(diff10)
d5 <-  density(diff5)
d3 <-  density(diff3)

plot(d10, xlim = c(-5, 1),
     ylim = c (0, 4),
     xlab = "",
     main = "", las = 1)
polygon(d10, col = tBlack)
lines(d5)
polygon(d5, col = tOrange)
lines(d3)
polygon(d3, col = tBlue)
mtext(side = 1,
      "standardized distance from archaeological sample\n (Z-score)", 
      outer = FALSE,
      line = 3.5)

legend("topleft",                                 
       legend = c("CV = 10%", 
                  "CV = 5%", 
                  "CV = 3%"), 
       fill = c(tBlack, tOrange, tBlue),     
       bty = "n" )                                

dev.off()
```

```{r}
knitr::include_graphics("figures/Garvey_Figure3b.pdf")
```


